---
title: "TSL Topic Analysis: Insights"
author: "Benji Lu, Kai Fukutaki, Ziqi Xiong"
date: "December 12, 2015"
output: ioslides_presentation
---
```{r, echo=FALSE, message=FALSE}
require(tidyr)
require(dplyr)
require(ggplot2)
require(lubridate)
require(gridExtra)
source('helpers.R')

# load articles
articles <- readData()
```
## Overall Articles Per Writer
```{r, echo=FALSE}
# get dataframe of number of articles written by each unique writer
articles_per_writer = articles %>%
  group_by(profile_id) %>%
  summarise(total = n())

# histogram of how many different pieces writers have written
articles_per_writer %>%
  ggplot(aes(x = total)) + geom_histogram(binwidth = 1, col=I('gray')) + ggtitle('Histogram of Article Count by Writer') + xlab('Number of Articles Written') + ylab('Number of Writers')
```

## Articles Per Writer by Section

```{r, echo=FALSE, fig.width=8.5, fig.height=6}
# important to note in the above that at least some of the counts include guest
# writers for opinions pieces who only write one or maybe a few pieces in their
# time in Claremont

# CHECK THE ORDERING OF THE SECTIONS TO MAKE SURE IT'S RIGHT
sections = c('News','LS','Opinions','Sports')
for (num in 1:4) {
  section <- sections[num]
  assign(paste(section, 'section_articles_per_writer', sep='_'), articles %>%
    filter(section_id == num) %>%
    group_by(profile_id) %>%
    summarise(total = n()) %>%
    ggplot(aes(x = total)) + geom_histogram(binwidth = 1, col=I('gray')) + ggtitle(section) + xlab('Number of Articles Written') + ylab('Number of Writers'))
}
grid.arrange(News_section_articles_per_writer,Opinions_section_articles_per_writer,LS_section_articles_per_writer,Sports_section_articles_per_writer, ncol=2)
```

## Top Writers

## Wes Haas
```{r, echo=FALSE, results='hide'}
# get top writer
sort(articles_per_writer$total)
row(articles_per_writer)[articles_per_writer[,2]==68]
articles_per_writer[69,]$profile_id
```
```{r, echo=FALSE}
# Wes Haas PO '14
(articles %>%
  filter(profile_id==81))[c(1:10),c(2,4)]
```

## Editorial Board {.smaller}
```{r, echo=FALSE, results='hide'}
# get top writer
row(articles_per_writer)[articles_per_writer[,2]==64]
articles_per_writer[34,]$profile_id
```
```{r, echo=FALSE}
# FIGURE OUT HOW TO MAKE PUBLISHED DATE TWO COLUMNS ON THE BOTTOM TO AVOID CUTOFF
(articles %>%
  filter(profile_id==39))[c(1:10),c(2,4)]
```

## Editorial Board {.smaller}
```{r, echo=FALSE, results='hide'}
# get top writer
row(articles_per_writer)[articles_per_writer[,2]==61]
articles_per_writer[148,]$profile_id
```
```{r echo=FALSE}
# FIGURE OUT HOW TO MAKE PUBLISHED DATE TWO COLUMNS ON THE BOTTOM TO AVOID CUTOFF
(articles %>%
  filter(profile_id==176))[c(1:10),c(2,4)]
```

## Tim Taylor {.smaller}
```{r, echo=FALSE, results='hide'}
# get top writer
row(articles_per_writer)[articles_per_writer[,2]==57]
articles_per_writer[88,]$profile_id
```
```{r, echo=FALSE}
# FIGURE OUT HOW TO MAKE PUBLISHED DATE TWO COLUMNS ON THE BOTTOM TO AVOID CUTOFF
(articles %>%
  filter(profile_id==103))[c(1:10),c(2,4)]
```

## The Student Life Staff {.smaller}
```{r, echo=FALSE, results='hide'}
# get top writer
row(articles_per_writer)[articles_per_writer[,2]==51]
articles_per_writer[147,]$profile_id
```
```{r, echo=FALSE}
# FIGURE OUT HOW TO MAKE PUBLISHED DATE TWO COLUMNS ON THE BOTTOM TO AVOID CUTOFF
(articles %>%
  filter(profile_id==175))[c(1:10),c(2,4)]
```

## Nachi Baru {.smaller}
```{r, echo=FALSE, results='hide'}
# get top writer
row(articles_per_writer)[articles_per_writer[,2]==48]
articles_per_writer[562,]$profile_id
```
```{r, echo=FALSE}
# FIGURE OUT HOW TO MAKE PUBLISHED DATE TWO COLUMNS ON THE BOTTOM TO AVOID CUTOFF
(articles %>%
  filter(profile_id==634))[c(1:10),c(2,4)]
```

## Retention Rate
```{r, echo=FALSE, results='hide'}
articles = articles %>%
  mutate(yearPublished = year(published_date)) %>%
  mutate(monthPublished = month(published_date)) %>%
  mutate(dayPublished = day(published_date)) %>%
  mutate(semester = ifelse(monthPublished < 6, 'spring', 'fall'))

retention = articles %>%
  group_by(profile_id) %>%
  mutate(maxYear = max(yearPublished),
         minYear = min(yearPublished),
         semesterOff = ifelse(maxYear==yearPublished & semester=='fall','fall','spring'),
         semesterJoin = ifelse(minYear==yearPublished & semester=='spring','spring','fall'),
         semestersTotal = ifelse(semesterOff==semesterJoin,(maxYear-minYear)*2+1, (maxYear-minYear)*2))

retention = retention[!duplicated(retention[,6]),]
```
```{r, echo=FALSE}
retention %>%
  group_by(semestersTotal) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = semestersTotal)) + geom_line(aes(y=n)) + geom_point(aes(y=n)) + xlab('Number of Semesters Writing for TSL') + ylab('Number of Writers') + ggtitle('Retention Rate of TSL Writers')
```

## Retention Rate by Section
```{r, echo=FALSE}
# FIX X-AXIS INTERVAL
retention %>%
  filter(section_id != 5) %>%
  group_by(semestersTotal, section_id) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = semestersTotal, y=n, color=factor(section_id))) + geom_point() + geom_line() + scale_color_discrete(name="Section", labels = c('News','Sports','Opinions','L&S')) + xlab('Number of Semesters Writing for TSL') + ylab('Number of Writers') + ggtitle('Retention Rate by Section')
```

## Retention Rate by Section (log)
```{r, echo=FALSE}
retention %>%
  filter(section_id != 5) %>%
  group_by(semestersTotal, section_id) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = log(semestersTotal), y=n, color=factor(section_id))) + geom_point() + geom_line() + scale_color_discrete(name="Section", labels = c('News','Sports','Opinions','L&S')) + xlab('log(Number of Semesters Writing for TSL)') + ylab('Number of Writers') + ggtitle('Retention Rate by Section')
```

## Article Views

## Most-Viewed Articles (Fall 2015) {.smaller}
```{r, echo=FALSE}
clicks <- read.csv('articles.csv')
clicks <- clicks[,c(1,4)]
names(clicks) <- c('id','clicks')

# most viewed articles (title, date published, writer, number of clicks)
(articles %>%
  left_join(clicks, by = 'id') %>%
  arrange(desc(clicks)) %>%
  slice(1:15))[,c(2,4,7,12)]
```

## Least-Viewed Articles (Fall 2015) {.smaller}
```{r, echo=FALSE}
clicks <- articles %>%
  filter(yearPublished == 2015 & semester == 'fall') %>%
  left_join(clicks, by = 'id')

# get info on least viewed pieces
(clicks %>%
  arrange(clicks) %>%
  slice(1:10))[,c(2,4,7,12)]

# interesting that they were all from the 12/4 issue. might be because
# it was the last issue and nothing really happened (right after Thanksgiving break).
```

## Least-Viewed Articles (Fall 2015) {.smaller}
```{r, echo=FALSE}
# exclude those from the data
(clicks %>%
  filter(published_date != as.Date('2015-12-04')) %>%
  arrange(clicks) %>%
  slice(1:10))[,c(2,4,7,12)]
```

## Summary of Article Views (Fall 2015)
```{r, echo=FALSE}
# plot distribution of views in fall 2015
clicks %>%
  ggplot(aes(x = semester, y = clicks)) + geom_boxplot() + xlab('Semester') + ylab('Cumulative Number of Clicks') + ggtitle('Distribution of Clicks of TSL Fall 2015 Articles')

# some huge outliers
```

## Summary of Article Views (Fall 2015)
```{r, echo=FALSE, warning=FALSE}
# exclude outliers from view
clicks %>%
  ggplot(aes(x = semester, y = clicks)) + geom_boxplot() + scale_y_continuous(limits = c(0,500)) + xlab('Semester') + ylab('Cumulative Number of Clicks') + ggtitle('Distribution of Clicks of TSL Fall 2015 Articles')
```

## Summary of Article Views by Section (Fall 2015)
```{r, echo=FALSE}
# do the same but by section
clicks %>%
  filter(section_id != 5) %>%
  mutate(section = ifelse(section_id==1, 'News', ifelse(section_id==2, 'Sports', ifelse(section_id==3, 'Opinions', 'L&S')))) %>%
  ggplot(aes(x = factor(section), y = clicks)) + geom_boxplot() + xlab('Section') + ylab('Cumulative Number of Clicks') + ggtitle('Distribution of Clicks by Section of TSL Fall 2015 Articles')
```

## Summary of Article Views by Section (Fall 2015)
```{r, echo=FALSE, warning=FALSE}
# exclude outliers from view
clicks %>%
  filter(section_id != 5) %>%
  mutate(section = ifelse(section_id==1, 'News', ifelse(section_id==2, 'Sports', ifelse(section_id==3, 'Opinions', 'L&S')))) %>%
  ggplot(aes(x = factor(section), y = clicks)) + geom_boxplot() + scale_y_continuous(limits = c(0,1250)) + xlab('Section') + ylab('Cumulative Number of Clicks') + ggtitle('Distribution of Clicks by Section of TSL Fall 2015 Articles')
```

## Most-Viewed Writers (Average) (Fall 2015)
```{r, echo=FALSE}
# we can try to identify the most clicked writers on average
(clicks %>%
  filter(section_id != 5) %>%
  group_by(author_name) %>%
  summarise(average_views = mean(clicks), articles_published=n()) %>%
  arrange(desc(average_views)) %>%
  slice(1:10))
```

## Most-Viewed Consistent Writers (Average) (Fall 2015)
```{r, echo=FALSE}
# a lot of the above are one-hit pieces. let's see among the more consistent writers
(clicks %>%
  filter(section_id != 5) %>%
  group_by(author_name) %>%
  summarise(average_views = mean(clicks), articles_published=n()) %>%
  filter(articles_published>3) %>%
  arrange(desc(average_views)) %>%
  slice(1:10))
```

## Least-Viewed Writers (Average) (Fall 2015)
```{r, echo=FALSE}
# same but for least clicked writers on average
(clicks %>%
  filter(section_id != 5) %>%
  group_by(author_name) %>%
  summarise(average_views = mean(clicks), articles_published=n()) %>%
  arrange(average_views) %>%
  slice(1:10))
```

## Least-Viewed Consistent Writers (Average) (Fall 2015)
```{r, echo=FALSE}
# again with repeat writers only
(clicks %>%
  filter(section_id != 5) %>%
  group_by(author_name) %>%
  summarise(average_views = mean(clicks), articles_published=n()) %>%
  filter(articles_published>3) %>%
  arrange(average_views) %>%
  slice(1:10))
```